---
title: "Hands-on Exercise 1"
author: "Yang Lu"
---

# Getting Started

## Installing and launching R packages

The code chunk below use p_load() of pacman package to check if tidyverse packages are installed in the computer. If they are , then they will be launched into R

```{r}
pacman::p_load(tidyverse)

```

## Importing the data

```{r}

exam_data <- read_csv("data/Exam_data.csv")
```

# Plotting a simple bar chart

```{r}
ggplot(data= exam_data,
       aes(x=RACE))+
  geom_bar()
```

```{r}
ggplot(data=exam_data,
       aes(x=MATHS))+
  geom_bar()
```

#Plotting a simple dot plot

```{r}
ggplot(data=exam_data,
       aes(x=MATHS))+
  geom_dotplot()

```

## Changing bin/binwith size

```{r}

ggplot(data=exam_data,
       aes(x=MATHS)) +
  geom_dotplot(dotsize=0.5)

```

## Y-axis is not very useful, using scale_y_continuous() to hide
```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot(binwidth=2.5,    #group scores into bins that are 2.5 units wide
               dotsize = 0.5) +      #dotsize=0.5 is half of default size
   scale_y_continuous(NULL,  #Hide the y-axis title
                     breaks = NULL)   #Hide Y-axis scales

```

#Plotting a simple histogram
```{r}
ggplot(data=exam_data,
       aes(x=MATHS))+
  geom_histogram()

```

### Modifying geometric object by changing geom (add fill="color")
```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20,            
                 color="black",      
                 fill="light blue")  
```

### Modifying geometric object by changing aes(add fill ="sub-group") 

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           fill = GENDER)) +
  geom_histogram(bins=20, 
                 color="grey30")

```

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           color= GENDER)) +
  geom_histogram(bins=20, 
                 color="grey30")
```
Note：different between using fill and colour 

fill controls inside fill  [Works on ： bar charts, histograms，boxplots etc.]
colour（color）controls border/outline color [Works on ： Line plots, scatter plots，edges etc.]


#Plotting a simple density

```{r}
ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_density()           
```
```{r}
ggplot(data=exam_data, 
       aes(x = MATHS, 
           colour = GENDER)) +
  geom_density()

```





```{r}
ggplot(data=exam_data, 
       aes(x = MATHS, 
           fill = GENDER)) +
  geom_density()

```


#Plotting a simple boxplot

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS,       
           x= GENDER)) +    
  geom_boxplot()            

```
##Using notch to visually compare the media
- if not overlap, means the medians are significantly different
- if overlap, means the difference between medians may not be statistically significant

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot(notch=TRUE)
```


#Plotting a simple violin

A violin plot is a data visualization that combines: boxplot (shows median and quartiles) + density plot (shows the distribution shape of the data)

The “width” of the graphic is indicated by
-Wider parts of the violin mean more data in that range
-Narrower parts mean fewer data there


```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_violin()

```

#Plotting a simple scatterplot

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point()            

```


#Combine the geom objects

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, 
           x= GENDER)) +
  geom_boxplot() +                    
  geom_point(position="jitter", #moves dots slightly left and right to avoid overlap
             size = 0.5)        
```


#Essential Grammatical Elements in ggplot2: stat
stat_() function and override the default geom
geom_() function and override the default stat




## Add means value by stat_summary() / geom() method
```{r}

ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  stat_summary(geom = "point",       
               fun = "mean",         
               colour ="red",        
               size=4)               


```
```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= GENDER)) +
  geom_boxplot() +
  geom_point(stat="summary",    #draw it as a dot     
             fun="mean",        #the value shown is the mean  
             colour="red",       # the dot colour   
             size=4)          # the dot size

```
