---
title: "Hands-on Exercise 2"
author: "Yang Lu"
Modified: "23rd April 2025"
---

# Lesson 2: Designing Graphs to Enlighten

```{r}
pacman::p_load(tidyverse)
```

## Installing and loading all required libraries

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse) 
```

## Importing data

```{r}
exam_data <- read_csv("Exam_data.csv")
```

# 2.3 Beyond ggplot2 Annotation: ggrepel

### As default \_overlapping text

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")

```

### Working with ggrepel

#### applying **geom_Label_repel** ：Aim of ggrepel is to repel overlapping text

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")


```

# 2.4 Beyond ggplot2 Themes

[Themes](https://ggplot2.tidyverse.org/reference/ggtheme.html)

::: panel-tabset
##theme_gray()

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20,   #The data is split into 20 bars 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores") 

```

## theme_bw()

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20,   #The data is split into 20 bars 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_bw() +
  ggtitle("Distribution of Maths scores") 

```

## theme_classic()

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20,   #The data is split into 20 bars 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_classic() +
  ggtitle("Distribution of Maths scores") 

```

## theme_dark()

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20,   #The data is split into 20 bars 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_dark() +
  ggtitle("Distribution of Maths scores")

```

## theme_light()

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20,   #The data is split into 20 bars 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_light() +
  ggtitle("Distribution of Maths scores")
```

## theme_linedraw()

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20,   #The data is split into 20 bars 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_linedraw() +
  ggtitle("Distribution of Maths scores")
```

## theme_minimal()

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20,   #The data is split into 20 bars 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_minimal() +
  ggtitle("Distribution of Maths scores")
```

## theme_void()

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20,   #The data is split into 20 bars 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_void() +
  ggtitle("Distribution of Maths scores")
```
:::

## 2.4.1 Working with ggtheme package

### Using Economist theme

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_economist()

```

It also provides some extra geoms and scales for ‘ggplot2’. Consult this vignette to learn more.

[this vignette](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/)

## 2.4.2 Working with hrbthems package

provides a base theme that focuses on typographic elements, including where various labels are placed as well as the fonts that are used.

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum()
```

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  ggtitle("Distribution of Maths scores") +
  theme_ipsum(axis_title_size = 18,
              base_size = 15,
              grid = "Y")

```

# 2.5 Beyond Single Graph

### Maths scores as p1

```{r}
p1 <- ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") + 
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of Maths scores")

print(p1) # the graphic save in p1, using print(p1) or directly write p1 to show 
```

### English scores as p2

```{r}
p2 <- ggplot(data=exam_data, 
             aes(x = ENGLISH)) +
  geom_histogram(bins=20, 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  coord_cartesian(xlim=c(0,100)) +
  ggtitle("Distribution of English scores")
print(p2)
```

### Drawing a scatterplot for English score versus Maths score

```{r}
p3 <- ggplot(data=exam_data, 
             aes(x= MATHS, 
                 y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")

p3
```

## 2.5.1 Creating Composite Graphics: pathwork methods

For combining several graphs using
