---
title: "Hands-on Exercise 2"
author: "Yang Lu"
---


Lesson 2: Designing Graphs to Enlighten



```{r}
pacman::p_load(tidyverse)
```

Installing and loading all required libraries

```{r}
pacman::p_load(ggrepel, patchwork, 
               ggthemes, hrbrthemes,
               tidyverse) 
```

Importing data 

```{r}
exam_data <- read_csv("Exam_data.csv")
```


2.3 Beyond ggplot2 Annotation: ggrepel

::: panel-tabset

##overlapping text

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label(aes(label = ID), 
             hjust = .5, 
             vjust = -.5) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")

```

## Working with ggrepel

applying geom_Label_repel ：Aim of ggrepel is to repel overlapping text 
```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, 
           y=ENGLISH)) +
  geom_point() +
  geom_smooth(method=lm, 
              size=0.5) +  
  geom_label_repel(aes(label = ID), 
                   fontface = "bold") +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  ggtitle("English scores versus Maths scores for Primary 3")


```
:::

2.4 Beyond ggplot2 Themes
comments:
 [Themes](https://ggplot2.tidyverse.org/reference/ggtheme.html)

::: panel-tabset

##theme_gray()

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20,   #The data is split into 20 bars 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_gray() +
  ggtitle("Distribution of Maths scores") 

```
## theme_bw()

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20,   #The data is split into 20 bars 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_bw() +
  ggtitle("Distribution of Maths scores") 

```

## theme_classic()
```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20,   #The data is split into 20 bars 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_classic() +
  ggtitle("Distribution of Maths scores") 

```

## theme_dark()
```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20,   #The data is split into 20 bars 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_dark() +
  ggtitle("Distribution of Maths scores")

```


## theme_light()
```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20,   #The data is split into 20 bars 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_light() +
  ggtitle("Distribution of Maths scores")
```

## theme_linedraw() 

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20,   #The data is split into 20 bars 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_linedraw() +
  ggtitle("Distribution of Maths scores")
```


## theme_minimal()

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20,   #The data is split into 20 bars 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_minimal() +
  ggtitle("Distribution of Maths scores")
```


## theme_void()

```{r}
ggplot(data=exam_data, 
             aes(x = MATHS)) +
  geom_histogram(bins=20,   #The data is split into 20 bars 
                 boundary = 100,
                 color="grey25", 
                 fill="grey90") +
  theme_void() +
  ggtitle("Distribution of Maths scores")
```


