---
title: "Take home Exercise 2"
author: "Yang Lu"
date: "29 May 2025"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true  
  echo: true  
  warning: false 
  freeze: true 
---

# VAST Challenge 2025-Mini-Challenge 3

For the exercise, our group will choose MC 3 as our main topic, the detail background referring to [VAST MC3](https://vast-challenge.github.io/2025/MC3.html)

## 1.0 Background

**Oceanus: From Fishing Nets to Tourism Tensions**

Once defined by its fishing traditions, Oceanus has transformed dramatically in the last decade. After a crackdown on illegal fishing, those involved redirected investments toward ocean tourism, creating community friction. When international star Sailor Shift announced plans to film locally, investigative journalist Clepper Jessen uncovered troubling connections around the temporary closure of Nemo Reef. His work exposed expedited approvals and secretive arrangements linking island officials, Sailor's team, influential local families, and conservationists The Green Guardians—suggesting a story of corruption hiding beneath paradise's surface.

### **2.0 Overall Tasks**

Clepper has built a knowledge graph from two weeks of intercepted radio communications in Oceanus. Here are the key tasks to be analyzed:

**Task 1: Temporal Communication Patterns**

1.  Create visual analytics to identify daily patterns in communication timing

2.  Analyze how these patterns changed over the two-week observation period

3.  Focus on a specific entity and use this information to determine who has influence over them.

**Task 2: Relationship Analysis**

1.  Develop visualizations to explore connections between vessels and people

2.  Identify closely associated groups and the predominant topics for each group

**Task 3: Pseudonym Identification**

1.  Expand visualizations to reveal who is using pseudonyms

2.  Visual analytics to identify common entity

3.  Explain how understanding pseudonyms changes interpretation of activities

## 3.0 Setting up

### 3.1 Loading packages

| Package | Description |
|:------------------|:----------------------------------------------------|
| jsonlite | JSON parsing and generation |
| tidyverse | Core data-science suite (dplyr, ggplot2, etc.) |
| SmartEDA | Automated exploratory data analysis |
| tidygraph | Tidy tools for graph-data manipulation and analysis |
| ggraph | Grammar-based graph/network visualization |
| lubridate | Easy parsing and manipulation of dates/times |
| dplyr | Grammar of data manipulation (filter, mutate, etc.) |
| patchwork | Combine multiple ggplots into a unified layout |
| purrr | Functional programming tools for lists and vectors |
| scales | Scale functions for visualization (axis breaks, formatting) |
| RColorBrewer | Color palettes for maps and plots |
| plotly | Interactive web-based charts |
| igraph | Network analysis and graph algorithms |
| viridis | Color scales for scientific visualization (colorblind-friendly) |
| forcats | Tools for working with categorical variables (factors) |
| knitr | Dynamic report generation and literate programming |
| visNetwork | Interactive network visualization in R |
| htmltools | Tools for HTML generation and dependency management |
| stringr | Consistent, simple tools for string manipulation |

```{r}
pacman::p_load(
  jsonlite,tidyverse,
  SmartEDA,tidygraph,
  ggraph,lubridate,
  dplyr,patchwork,
  purrr,scales,
  RColorBrewer,plotly,
  igraph,viridis,
  forcats,knitr,
  visNetwork,htmltools,
  stringr)
```

### 3.2 Loading data

```{r}
MC3_graph <- fromJSON("data/MC3_graph.json")
glimpse(MC3_graph)
```

```{r}
MC3_schema <- fromJSON("data/MC3_schema.json") 
glimpse(MC3_schema)
```

### 3.3 Examining the data structure

```{r}
nodes_tbl <- as_tibble(MC3_graph$nodes)
edges_tbl <- as_tibble(MC3_graph$edges)
```

```{r}
kable(head(nodes_tbl, 10))
kable(head(edges_tbl, 10))


```

after split to nodes and edges, the data show **3226 edges** and **1159 nodes**

## 4.Initial EDA

It is time for us to apply appropriate EDA methods to examine the data.

In the code chunk below, ExpCatViz() of SmartEDA package is used to reveal the frequency distribution of all categorical fields in nodes_tbl tibble dataframe.

```{r}
ExpCatViz(data = nodes_tbl,
          col = "red")
```

On the other hands, code chunk below uses ExpCATViz() of SmartEDA package to reveal the frequency distribution of all categorical fields in MC3_edges tibble data frame.

```{r}
ExpCatViz(data=edges_tbl,
          col="lightblue")


```

## 5.0 Task 1-Clepper found that messages frequently came in at around the same time each day.

### Task 1.1: Create visual analytics to identify daily patterns in communication timing

```{r}
ggplot(data=nodes_tbl,
      aes(y=`type`))+
  geom_bar()


```

```{r}
ggplot(data=nodes_tbl,
      aes(y=`sub_type`))+
  geom_bar()

```

#### **Step 1: Generate Node Sub-Type Inventory**

```{r}

# Extract unique sub_types for each node type
entities      <- sort(unique(nodes_tbl$sub_type[nodes_tbl$type == "Entity"]))
events        <- sort(unique(nodes_tbl$sub_type[nodes_tbl$type == "Event"]))
relationships <- sort(unique(nodes_tbl$sub_type[nodes_tbl$type == "Relationship"]))

node_inventory <- list(
  Entities      = entities,
  Events        = events,
  Relationships = relationships
)


```

#### **Step 2: List down type and relevant sub types**

```{r}
for (grp in names(node_inventory)) {
  cat("**", grp, "** (", length(node_inventory[[grp]]), "):\n", sep = "")
  cat(paste0("  - ", node_inventory[[grp]]), sep = "\n")
  cat("\n")
}



```

#### Step 3: Create communication nodes

based on the above list, we noticed that based on the task requirement, now we only focus on type="event" and sub_type ="communication"

```{r}
comm_nodes <-nodes_tbl %>%
  filter(type=="Event",
         sub_type == "Communication") %>%
  select(event_id = id,timestamp)

glimpse(comm_nodes)
```

#### **Step 4: Define sender and receiver**

```{r}
ggplot(data=edges_tbl,
      aes(y=`type`))+
  geom_bar()

```

#### **Step 5: Extract data from sender and receiver**

```{r}
sent_edges <-edges_tbl %>% filter(type=="sent")%>%
  select(sender = source,event_id=target)

recv_edges <-edges_tbl %>% filter(type=="received")%>%
  select(event_id = source,receiver=target)
```

#### **Step 6: using inner_join to combine sender and receiver**

```{r}
msgs <- sent_edges %>%
  inner_join(recv_edges, by="event_id") %>%
  inner_join(comm_nodes, by="event_id") %>%
  mutate(ts = ymd_hms(timestamp, tz="UTC")) %>%
  mutate(date = as_date(ts), hour = hour(ts))

```

#### **Step 7: Visualizing the heat map**

```{r}
heatmap_data <- msgs %>%
  count(date, hour) %>%
  complete(
    date = seq(min(date), max(date), by = "1 day"),
    hour = 0:23,
    fill = list(n = 0)
  )

ggplot(heatmap_data, aes(x=hour, y=date, fill=n)) +
  geom_tile(color="white") +
  scale_x_continuous(
    breaks = 0:23,
    expand = c(0,0)
  )+
  scale_y_date(
    breaks = seq(min(heatmap_data$date), max(heatmap_data$date), by = "1 day"),
    date_labels = "%b %d",
    expand = c(0, 0)
  ) +
 scale_fill_distiller(
    name    = "Messages",
    palette = "Spectral",
    direction = 1
  ) +
  
  labs(
    title = "Daily Communication Patterns",
    x     = "Hour of Day",
    y     = "Date"
  ) +
  theme_minimal() +
  theme(
    axis.text.x  = element_text(hjust = 0.5),
    axis.text.y  = element_text(size = 8),
    panel.grid  = element_blank()
  )


```

### Task 1.2 Compare the shape of daily rhythm between week 1 and week 2

```{r,warning=FALSE}
msgs <- msgs %>%
  mutate(
    week = if_else(date <= min(date) + days(6), "Week 1", "Week 2")
  )
# Count and normalize within each week
week_patterns <- msgs %>%
  count(week, hour) %>%
  group_by(week) %>%
  mutate(proportion = n / sum(n))


hour_breaks <- seq(min(week_patterns$hour), max(week_patterns$hour), by = 1)
prop_breaks <- seq(
  0,
  ceiling(max(week_patterns$proportion) * 100) / 100,
  by = 0.02
)

ggplot(week_patterns, aes(x = hour, y = proportion, color = week)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_x_continuous(breaks = hour_breaks) +
  scale_y_continuous(
    breaks = prop_breaks,
    labels = percent_format(accuracy = 1)
  ) +
  labs(
    title = "Hourly Communication Patterns: Week 1 vs Week 2",
    x     = "Hour of Day",
    y     = "Percent of Total Messages"
  ) +
  theme_light(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 16, hjust = 0.5),
    legend.position = "top",
    legend.title    = element_blank(),
    panel.grid.major = element_line(color = "grey80"),
    panel.grid.minor = element_line(color = "grey90"),
    axis.text.x     = element_text(vjust = 0.5),
    axis.text.y     = element_text(size = 8)
  )



```

::: {.callout-caution style="background-color: white;" icon="false"}
## Conclusion

The visualization of hourly communication patterns reveals that message activity in Oceanus follows a pronounced daily cycle, with distinct peaks and lows across both observed weeks.

In Week 1, communication is highly concentrated in the morning hours, particularly between 8 AM and 12 PM, where a majority of the messages are exchanged.

By contrast, Week 2 shows a broader distribution of communication throughout the day, with notable increases in afternoon and evening activity.

This shift suggests that while the first week’s communications were more focused and possibly related to regular planning or operational updates, the second week’s patterns reflect heightened activity or more dynamic coordination, potentially due to emergent events or increased urgency among entities.
:::

### Task 1.3: Focus on a specific entity and use this information to determine who has influence over them.

```{r}
entity_info <- nodes_tbl %>%
  filter(type == "Entity") %>%
  select(id, target_subtype = sub_type)

msgs2 <- msgs %>%
  left_join(entity_info, by = c("receiver" = "id"))


top_inf <- msgs2 %>%
  filter(!is.na(receiver)) %>%
  count(target_subtype, receiver, sender, sort = TRUE) %>%
  group_by(target_subtype, receiver) %>%
  slice_max(n, n = 10) %>%
  ungroup()

plot_list <- top_inf %>%
  group_split(target_subtype) %>%
  map(~{
    subtype <- unique(.x$target_subtype)
    ggplot(.x, aes(
        x     = reorder(sender, n, .desc=TRUE),
        y     = n,
        fill  = receiver,
        text  = paste0(
          "Subtype: ", subtype, "<br>",
          "Sender: ", sender, "<br>",
          "Receiver: ", receiver, "<br>",
          "Count: ", n
        )
      )) +
      geom_col(color = "white") +                    
      scale_fill_viridis_d(option = "turbo", name="Receiver") +
      coord_flip() +
      labs(
        title = paste("Subtype:", subtype),
        x     = "Influencer (Sender)",
        y     = "Message Count"
      ) +
      theme_minimal(base_size = 12) +
      theme(
        strip.text       = element_text(face = "bold"),
        axis.text.y      = element_text(size = 8),
        panel.grid.major = element_line(color = "grey80"),
        panel.grid.minor = element_line(color = "grey90"),
        legend.position  = "bottom"
      )
  })


plots <- map(plot_list, ~ggplotly(.x, tooltip = "text"))
tagList(plots)


```

::: panel-tabset
### Subtype:Group

```{r,warning=FALSE}

ee_edges <- msgs %>% select(from = sender, to = receiver)
entity_nodes <- nodes_tbl %>%
  filter(type == "Entity") %>%
  select(id, name)

entity_graph <- tbl_graph(nodes = entity_nodes, edges = ee_edges, directed = TRUE)


target <- "Neptune"


g_ig <- as.igraph(entity_graph, strict = FALSE)


root_idx <- which(V(g_ig)$name == target)


ego_vs <- unique(c(
  root_idx,
  neighbors(g_ig, root_idx, mode = "in"),
  neighbors(g_ig, root_idx, mode = "out")
))


subg <- induced_subgraph(g_ig, vids = ego_vs)


ego_graph <- as_tbl_graph(subg)


ggraph(ego_graph, layout = "kk") +
  geom_edge_link(
    aes(label = NULL),
    arrow = arrow(length = unit(4, "mm")),
    end_cap = circle(3, 'mm'),
    color = "grey70"
  ) +
  geom_node_point(
    aes(filter = (name == target)),
    size  = 6,
    color = "firebrick"
  ) +
  geom_node_point(
    aes(filter = (name != target)),
    size  = 4,
    color = "steelblue"
  ) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  labs(title = paste("Influence Network of", target)) +
  theme_graph()
```

### Subtype: Location

```{r,warning=FALSE}

ee_edges <- msgs %>% select(from = sender, to = receiver)
entity_nodes <- nodes_tbl %>%
  filter(type == "Entity") %>%
  select(id, name)

entity_graph <- tbl_graph(nodes = entity_nodes, edges = ee_edges, directed = TRUE)


target <- "Oceanus City Council"


g_ig <- as.igraph(entity_graph, strict = FALSE)


root_idx <- which(V(g_ig)$name == target)


ego_vs <- unique(c(
  root_idx,
  neighbors(g_ig, root_idx, mode = "in"),
  neighbors(g_ig, root_idx, mode = "out")
))


subg <- induced_subgraph(g_ig, vids = ego_vs)


ego_graph <- as_tbl_graph(subg)


ggraph(ego_graph, layout = "kk") +
  geom_edge_link(
    aes(label = NULL),
    arrow = arrow(length = unit(4, "mm")),
    end_cap = circle(3, 'mm'),
    color = "grey70"
  ) +
  geom_node_point(
    aes(filter = (name == target)),
    size  = 6,
    color = "firebrick"
  ) +
  geom_node_point(
    aes(filter = (name != target)),
    size  = 4,
    color = "steelblue"
  ) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  labs(title = paste("Influence Network of", target)) +
  theme_graph()
```

### Subtype:Organization

```{r,warning=FALSE}

ee_edges <- msgs %>% select(from = sender, to = receiver)
entity_nodes <- nodes_tbl %>%
  filter(type == "Entity") %>%
  select(id, name)

entity_graph <- tbl_graph(nodes = entity_nodes, edges = ee_edges, directed = TRUE)


target <- "Green Guardians"


g_ig <- as.igraph(entity_graph, strict = FALSE)


root_idx <- which(V(g_ig)$name == target)


ego_vs <- unique(c(
  root_idx,
  neighbors(g_ig, root_idx, mode = "in"),
  neighbors(g_ig, root_idx, mode = "out")
))


subg <- induced_subgraph(g_ig, vids = ego_vs)


ego_graph <- as_tbl_graph(subg)


ggraph(ego_graph, layout = "kk") +
  geom_edge_link(
    aes(label = NULL),
    arrow = arrow(length = unit(4, "mm")),
    end_cap = circle(3, 'mm'),
    color = "grey70"
  ) +
  geom_node_point(
    aes(filter = (name == target)),
    size  = 6,
    color = "firebrick"
  ) +
  geom_node_point(
    aes(filter = (name != target)),
    size  = 4,
    color = "steelblue"
  ) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  labs(title = paste("Influence Network of", target)) +
  theme_graph()
```

### Subtype:Person

```{r,warning=FALSE}

ee_edges <- msgs %>% select(from = sender, to = receiver)
entity_nodes <- nodes_tbl %>%
  filter(type == "Entity") %>%
  select(id, name)

entity_graph <- tbl_graph(nodes = entity_nodes, edges = ee_edges, directed = TRUE)


target <- "The Intern"


g_ig <- as.igraph(entity_graph, strict = FALSE)


root_idx <- which(V(g_ig)$name == target)


ego_vs <- unique(c(
  root_idx,
  neighbors(g_ig, root_idx, mode = "in"),
  neighbors(g_ig, root_idx, mode = "out")
))


subg <- induced_subgraph(g_ig, vids = ego_vs)


ego_graph <- as_tbl_graph(subg)


ggraph(ego_graph, layout = "kk") +
  geom_edge_link(
    aes(label = NULL),
    arrow = arrow(length = unit(4, "mm")),
    end_cap = circle(3, 'mm'),
    color = "grey70"
  ) +
  geom_node_point(
    aes(filter = (name == target)),
    size  = 6,
    color = "firebrick"
  ) +
  geom_node_point(
    aes(filter = (name != target)),
    size  = 4,
    color = "steelblue"
  ) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  labs(title = paste("Influence Network of", target)) +
  theme_graph()

```

### Vessel

```{r,warning=FALSE}

ee_edges <- msgs %>% select(from = sender, to = receiver)
entity_nodes <- nodes_tbl %>%
  filter(type == "Entity") %>%
  select(id, name)

entity_graph <- tbl_graph(nodes = entity_nodes, edges = ee_edges, directed = TRUE)


target <- "Reef Guardian"


g_ig <- as.igraph(entity_graph, strict = FALSE)


root_idx <- which(V(g_ig)$name == target)


ego_vs <- unique(c(
  root_idx,
  neighbors(g_ig, root_idx, mode = "in"),
  neighbors(g_ig, root_idx, mode = "out")
))


subg <- induced_subgraph(g_ig, vids = ego_vs)


ego_graph <- as_tbl_graph(subg)


ggraph(ego_graph, layout = "kk") +
  geom_edge_link(
    aes(label = NULL),
    arrow = arrow(length = unit(4, "mm")),
    end_cap = circle(3, 'mm'),
    color = "grey70"
  ) +
  geom_node_point(
    aes(filter = (name == target)),
    size  = 6,
    color = "firebrick"
  ) +
  geom_node_point(
    aes(filter = (name != target)),
    size  = 4,
    color = "steelblue"
  ) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  labs(title = paste("Influence Network of", target)) +
  theme_graph()
```
:::

## Task 2: Clepper has noticed that people often communicate with (or about) the same people or vessels, and that grouping them together may help with the investigation.

### Task 2.1 :Develop visualizations to explore connections between vessels and people

#### Step1:Clean the nodes and edges data

clean the raw nodes and edges data to ensure there are no missing values, all IDs are unique, and types are consistent.

```{r}

mc3_nodes_cleaned <- nodes_tbl %>%
  mutate(id = as.character(id)) %>%
  filter(!is.na(id)) %>%
  distinct(id, .keep_all = TRUE) %>%
  select(-thing_collected,everything())

mc3_edges_cleaned <- edges_tbl %>%
  rename(from_id = source, to_id = target) %>%
  mutate(across(c(from_id, to_id), as.character)) %>%
  filter(from_id %in% mc3_nodes_cleaned$id, to_id %in% mc3_nodes_cleaned$id) %>%
  filter(!is.na(from_id), !is.na(to_id))


```

#### Step 2: Extract person–vessel indirect links

Identify "Person" and "Vessel" nodes, and link them through "Relationship" nodes to extract all person–relationship–vessel triples.

```{r}
people_nodes <- mc3_nodes_cleaned %>% filter(sub_type == "Person")
vessel_nodes <- mc3_nodes_cleaned %>% filter(sub_type == "Vessel")
relationship_nodes <- mc3_nodes_cleaned %>% 
  filter(grepl("^Relationship_", id))

person_relationship_edges <- mc3_edges_cleaned %>%
  filter(from_id %in% people_nodes$id & to_id %in% relationship_nodes$id)

relationship_vessel_edges <- mc3_edges_cleaned %>%
  filter(from_id %in% relationship_nodes$id & to_id %in% vessel_nodes$id)

person_to_relationship <- person_relationship_edges %>%
  select(person_id = from_id, relationship_id = to_id)
relationship_to_vessel <- relationship_vessel_edges %>%
  select(relationship_id = from_id, vessel_id = to_id)

person_vessel_links <- person_to_relationship %>%
  inner_join(relationship_to_vessel, by = "relationship_id") %>%
  select(person_id, vessel_id, relationship_id)

```

#### Step 3: Add labels and types for nodes and edges

Add labels to each person and vessel node, and add relationship type to edges for visualization.

```{r}

person_vessel_links <- person_vessel_links %>%
  left_join(people_nodes %>% select(id, person_label = label), by = c("person_id" = "id")) %>%
  left_join(vessel_nodes %>% select(id, vessel_label = label), by = c("vessel_id" = "id"))

person_vessel_links <- person_vessel_links %>%
  mutate(relationship_type = stringr::str_extract(relationship_id, "(?<=Relationship_)[^_]+"))

edges_pv <- person_vessel_links %>%
  mutate(
    from = person_label,
    to = vessel_label
  ) %>%
  select(from, to, relationship_type)


```

#### Step 4: Prepare nodes/edges table for visualization

Combine all nodes (person + vessel) into a nodes dataframe with type and color, and add color to the edge dataframe.

```{r}
nodes <- data.frame(
  id = unique(c(edges_pv$from, edges_pv$to)),
  label = unique(c(edges_pv$from, edges_pv$to))
) %>%
  left_join(
    bind_rows(
      people_nodes %>% select(label, type = sub_type),
      vessel_nodes %>% select(label, type = sub_type)
    ),
    by = "label"
  ) %>%
  mutate(
    shape = ifelse(type == "Person", 16, 17),
    color = ifelse(type == "Person", "#FF8888", "#44C6C6")
  ) %>%
  select(id, label, type, shape, color)

relation_types <- sort(unique(edges_pv$relationship_type))
if (length(relation_types) <= 8) {
  relation_colors <- RColorBrewer::brewer.pal(length(relation_types), "Dark2")
} else {
  relation_colors <- colorRampPalette(RColorBrewer::brewer.pal(8, "Dark2"))(length(relation_types))
}
relation_color_map <- setNames(relation_colors, relation_types)

edges <- edges_pv %>%
  mutate(
    relationship = relationship_type,
    color = relation_color_map[relationship_type]
  ) %>%
  select(from, to, relationship, color)

```

#### Step 5: Plot static network with node/edge legends

Use ggraph to plot a static network diagram. Node types (person/vessel) have different shapes and colors, and the legend on the right automatically distinguishes them. Edge colors represent relationship types.

```{r}
graph_pv <- tbl_graph(
  nodes = nodes,
  edges = edges %>% rename(label = relationship),
  directed = TRUE
)

ggraph(graph_pv, layout = 'fr') +
  geom_edge_link(
    aes(color = label),
    arrow = arrow(length = unit(3, 'mm')),
    end_cap = circle(3, 'mm'),
    width = 1.4,
    alpha = 0.85,
    show.legend = TRUE
  ) +
  geom_node_point(
    aes(shape = type, color = type),
    size = 6,
    show.legend = TRUE
  ) +
  geom_node_text(
    aes(label = label),
    repel = TRUE,
    size = 3.8,
    vjust = 1.5,
    segment.color = NA 
  ) +
  scale_shape_manual(
    values = c("Person" = 16, "Vessel" = 17),
    name = "Node Type"
  ) +
  scale_color_manual(
    values = c("Person" = "#FF8888", "Vessel" = "#44C6C6"),
    name = "Node Type"
  ) +
  scale_edge_color_manual(
    values = relation_color_map,
    name = "Relationship"
  ) +
  guides(
    color = guide_legend(override.aes = list(size = 6)),
    shape = guide_legend(override.aes = list(size = 6)),
    edge_color = guide_legend(override.aes = list(linetype = 1, size = 2))
  ) +
  theme_void() +
  labs(title = "People–Vessel Interaction Network") +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold")
  )

 

```

### Task 2.2 ：Identify closely associated groups and the predominant topics for each group

#### Step 1: Data preparation

```{r}
# Convert directed graph to undirected for community detection
ig_undir <- as.undirected(as.igraph(graph_pv), mode = "collapse")

#run Louvain community detection
com <- cluster_louvain(ig_undir)
membership_vec <- membership(com) 

# assign community back to tbl_graph
graph_pv <- graph_pv %>% 
  activate(nodes) %>%
  mutate(community = as.factor(membership_vec[seq_len(n())]))

#extract node table with community info
nodes <- as_tibble(graph_pv, "nodes")


```

#### Step 2: Community membership summary

```{r}
group_summary <- nodes %>%
  group_by(community, type) %>%
  summarise(
    members = paste(label, collapse = ", "),
    .groups = "drop"
  )

kable(group_summary)


```

#### Step 3:Visualization of community structure

```{r}
n_communities <- length(unique(na.omit(nodes$community)))
if(n_communities > 8) {
  palette <- colorRampPalette(brewer.pal(8, "Set2"))(n_communities)
} else {
  palette <- brewer.pal(max(3, n_communities), "Set2")
}

relation_types <- unique(edges$label)
n_rel <- length(relation_types)
relation_colors <- brewer.pal(max(3, n_rel), "Dark2")
relation_color_map <- setNames(relation_colors, relation_types)

ggraph(graph_pv, layout = 'fr') +
  geom_edge_link(
    aes(color = label),
    arrow = arrow(length = unit(3, 'mm')), 
    end_cap = circle(3, 'mm'),
    width = 1.3,
    alpha = 0.9,
    show.legend = TRUE
  ) +
  geom_node_point(
    aes(shape = type, fill = community),
    size = 9,
    color = "black",
    stroke = 1.1,
    alpha = 1,
    show.legend = TRUE
  ) +
  geom_node_text(
    aes(label = label),
    repel = TRUE,
    size = 3,
    segment.color = NA
  ) +
  scale_shape_manual(
    values = c("Person" = 21, "Vessel" = 24),
    name = "Node Type"
  ) +
  scale_fill_manual(
    values = palette,
    name = "Community"
  ) +
  scale_edge_color_manual(
    values = relation_color_map,
    name = "Relationship"
  ) +
  guides(
    fill = guide_legend(
      override.aes = list(shape = 21, size = 7, color = "black", stroke = 1.2)
    ),
    shape = guide_legend(
      override.aes = list(size = 7, fill = "#CCCCCC", color = "black", stroke = 1.2)
    ),
    edge_color = guide_legend(
      override.aes = list(linetype = 1, size = 3, alpha = 1)
    )
  ) +
  theme_void() +
  labs(title = "People–Vessel Network: Community Structure") +
  theme(
    legend.position = "right",
    legend.box = "vertical",
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
  )
  



```

```{r}
group_summary$topic <- case_when(
  grepl("EcoVigil|Green Guardians|Nadia Conti", group_summary$members, ignore.case = TRUE) ~ "Environmentalism",
  grepl("Mako|Marlin|Neptune|Remora", group_summary$members, ignore.case = TRUE) ~ "Fishing/Leisure",
  grepl("Clepper Jensen|Knowles", group_summary$members, ignore.case = TRUE) ~ "Media/Investigation",
  grepl("Boss|Small Fry", group_summary$members, ignore.case = TRUE) ~ "Pseudonym Group",
  TRUE ~ "Other"
)

kable(group_summary)

```

::: {.callout-caution style="background-color: white;" icon="false"}
## Conclusion

The community detection analysis of the people–vessel interaction network revealed several closely associated groups within the Oceanus scenario. Each community, identified through the Louvain algorithm and confirmed by network visualization, displays distinct thematic characteristics based on the key members present.

-   One community centers around Nadia Conti and EcoVigil, strongly suggesting an environmentalist coalition, possibly focused on conservation activities around Nemo Reef.

-   Another group is dominated by traditional vessel names such as Mako, Marlin, and Neptune, along with key figures like Davis and Mrs. Money, indicating a cluster of actors involved in fishing and leisure vessel operations.

-   A smaller, tightly-knit group including Clepper Jensen and Knowles likely represents media or investigative interests, reflecting the narrative’s focus on journalistic inquiry.

-   Additionally, a community featuring members with pseudonymous names such as Boss and Small Fry appears to be engaged in covert or less formal operations.

These findings highlight the presence of distinct interest groups within the Oceanus community, each aligned around environmentalism, fishing/leisure activities, media/investigation, or covert communications. The clear separation and internal cohesion of these groups underscore the complex social dynamics at play, supporting the investigation into tensions and alliances underlying recent events in Oceanus.
:::

### **Task 3: Pseudonym Identification**

### Task 3.1 Expand visualizations to reveal who is using pseudonyms

To detect the pseudonyms, below code is based on two rules for `sub_type = "Person"`:

1.  Name starting with "The" :These are very likely pseudonyms, such as “The Lookout” or “The Intern”

2.  Obvious pseudonym keywords: Names that contain certain keywords commonly used as roles or code names (e.g., “Boss”, “Accountant”, “Middleman”, etc.).

```{r}
nodes_person_vessel <- nodes_tbl %>%
  filter(sub_type %in% c("Person","Vessel"))

pseudonym_keywords <- c("Boss", "Accountant", "Middleman", "Lookout", "Money", "Intern", "Small Fry")

nodes_person_vessel <- nodes_person_vessel %>%
  mutate(
    is_the_prefix = str_detect(label, "^The "),
    is_keyword = str_detect(label, paste(pseudonym_keywords, collapse = "|")),
    is_pseudonym = is_the_prefix | is_keyword
  )

pseudo_table <- nodes_person_vessel%>%
  filter(is_pseudonym == TRUE) %>%
  select(id, label, sub_type)

kable(pseudo_table)


```

::: {.callout-caution style="background-color: white;" icon="false"}
## Conclusion

By applying a rule-based pseudonym detection algorithm to nodes classified as sub_type = "Person", we identified entities whose labels either begin with "The" or match a curated set of known pseudonym keywords (e.g., “Boss”, “Accountant”, “Middleman”, etc.).

This approach systematically flagged a total of **seven** unique person nodes: "The Intern", "The Lookout", "The Accountant", "Mrs. Money", "The Middleman", "Boss", and "Small Fry".

These entities are highly likely to represent pseudonymous actors within the communication network, potentially serving as proxies or aliases to conceal the true identity or organizational role of the underlying individual.

The technical filtering strategy ensures a reproducible and objective process for pseudonym identification, which will be leveraged in downstream analyses of network influence, alias mapping, and the detection of covert activity within the Oceanus knowledge graph.
:::

### Task 3.2 ：Visual analytics to identify common entity (Pseudonym and their network)

```{r}
pseudo_labels <- pseudo_table$label
core_ids <- nodes_tbl %>% filter(sub_type %in% c("Person", "Vessel")) %>% pull(label)

pseudo_to_rel <- edges_tbl %>%
  filter(source %in% pseudo_labels & grepl("^Relationship_", target)) %>%
  select(pseudo = source, relationship = target, type_pseudo = type)

rel_to_person <- edges_tbl %>%
  filter(grepl("^Relationship_", source) & target %in% core_ids) %>%
  select(relationship = source, person = target, type_person = type)

pseudo_rel_network <- pseudo_to_rel %>%
  inner_join(rel_to_person, by = "relationship") %>%
  select(pseudonym = pseudo, relationship, person, type_pseudonym = type_pseudo, type_person)


extract_reltype <- function(x) {
  m <- regmatches(x, regexpr("_(.*?)_", x))
  if (length(m) == 0) return("Other")
  gsub("_", "", m)
}
rel_type_vec <- sapply(pseudo_rel_network$relationship, extract_reltype)
rel_types <- unique(rel_type_vec)
rel_colors <- setNames(RColorBrewer::brewer.pal(max(3, length(rel_types)), "Set2"), rel_types)


person_vessel_info <- nodes_tbl %>%
  filter(label %in% core_ids | label %in% pseudo_labels) %>%
  select(label, sub_type) %>%
  distinct(label, .keep_all = TRUE)

nodes_all <- unique(c(pseudo_rel_network$pseudonym, pseudo_rel_network$relationship, pseudo_rel_network$person))
nodes_viz <- tibble(id = nodes_all) %>%
  left_join(person_vessel_info, by = c("id" = "label"))

nodes_viz <- tibble(id = nodes_all) %>%
  left_join(person_vessel_info, by = c("id" = "label")) %>%
  mutate(
    group = ifelse(id %in% pseudo_labels, "Pseudonym",
                   ifelse(!is.na(sub_type) & sub_type == "Person", "Person",
                          ifelse(!is.na(sub_type) & sub_type == "Vessel", "Vessel",
                                 ifelse(grepl("^Relationship_", id), extract_reltype(id), "Other")))),
    label = ifelse(grepl("^Relationship_", id), "", id),
    title = ifelse(group == "Pseudonym", paste0("Pseudonym: ", id, "<br>Type: ", sub_type),
                   ifelse(group == "Person", paste0("Person: ", id),
                          ifelse(group == "Vessel", paste0("Vessel: ", id),
                                 ifelse(grepl("^Relationship_", id), paste0("Relationship: ", extract_reltype(id)), id)))),
    color = ifelse(group == "Pseudonym", "red",
                   ifelse(group == "Person", "skyblue",
                          ifelse(group == "Vessel", "orange",
                                 ifelse(group %in% rel_types, rel_colors[group], "grey")))),
    shape = ifelse(group == "Pseudonym", "square",
                   ifelse(group == "Person", "dot",
                          ifelse(group == "Vessel", "triangle",
                                 ifelse(group %in% rel_types, "diamond", "dot"))))
  )

edges_viz <- bind_rows(
  pseudo_rel_network %>%
    mutate(rel_type = extract_reltype(relationship)) %>%
    transmute(
      from = pseudonym, to = relationship,
      color = rel_colors[rel_type],
      arrows = "to", title = paste0("Pseudonym to Relationship: ", rel_type)
    ),
  pseudo_rel_network %>%
    mutate(rel_type = extract_reltype(relationship)) %>%
    transmute(
      from = relationship, to = person,
      color = rel_colors[rel_type],
      arrows = "to", title = paste0("Relationship to Person/Vessel: ", rel_type)
    )
)

all_legend <- list(
    list(label = "Pseudonym", shape = "square", color = "red"),
    list(label = "Person", shape = "dot", color = "skyblue"),
    list(label = "Vessel", shape = "triangle", color = "orange")
)

visNetwork(nodes_viz, edges_viz, width = "1600px", height = "900px") %>%
  visEdges(arrows = "to",smooth = FALSE) %>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE, selectedBy = "group") %>%
  visLegend(
    addNodes = all_legend,
    useGroups = FALSE
  ) %>%
  visPhysics(stabilization = TRUE) %>%
  visInteraction(
    dragNodes = TRUE,
    dragView = TRUE,
    zoomView = TRUE
  )



```

### Task 3.3 :How does your understanding of activities change given your understanding of pseudonyms?

::: {.callout-caution style="background-color: white;" icon="false"}
## Conclusion

With the explicit identification of pseudonyms such as “Boss,” “The Intern,” “The Accountant,” “Mrs. Money,” “The Middleman,” “The Lookout,” and “Small Fry,” my understanding of the activities within the Oceanus network changed significantly.

**Initially**, communications and relationships appeared to be fragmented among many individuals and vessels. However, after mapping pseudonyms onto the network, several important patterns emerged:

-   **Pseudonyms as Key Hubs:** In the visual network, the detected pseudonyms are not peripheral actors—they consistently occupy central hub positions, connecting to multiple vessels and people through various types of relationships (such as "Reports," "Coordinates," "Operates," and "Suspicious").

-   **Obscured True Roles:** What previously looked like many unrelated participants was actually a small set of influential actors operating under multiple pseudonyms. For example, “The Intern” and “Boss” are both connected to several vessels via "Coordinates" and "Operates" relationships, indicating they may be orchestrating or managing operations behind the scenes.

-   **Concealment of Illicit Activity:** The structure revealed that pseudonyms are often used to mask direct links between real individuals and vessels—particularly in suspicious or high-frequency communication events. This suggests intentional obfuscation, possibly to hide illicit coordination or the flow of sensitive information.

-   **Redefining Influence and Risk:** By highlighting pseudonyms, it becomes clear that influence within the network is far more centralized than the real-name view suggested. A few pseudonyms control or interact with a large proportion of vessels and people, making them high-value targets for further investigation.
:::
