---
title: "Take home Exercise 3"
author: "Yang Lu"
date: "15 May 2025"
date-modified: "last-modified"
format: html
editor: visual
execute: 
  eval: true  
  echo: true  
  warning: false 
  freeze: true 
---

# VAST Challenge 2025-Mini-Challenge 3

For the exercise, our group will choose MC 3 as our main topic, the detail background referring to [VAST MC3](https://vast-challenge.github.io/2025/MC3.html)

## 1.0 Background

**Oceanus: From Fishing Nets to Tourism Tensions**

Once defined by its fishing traditions, Oceanus has transformed dramatically in the last decade. After a crackdown on illegal fishing, those involved redirected investments toward ocean tourism, creating community friction. When international star Sailor Shift announced plans to film locally, investigative journalist Clepper Jessen uncovered troubling connections around the temporary closure of Nemo Reef. His work exposed expedited approvals and secretive arrangements linking island officials, Sailor's team, influential local families, and conservationists The Green Guardians—suggesting a story of corruption hiding beneath paradise's surface.

### **2.0 Overall Tasks**

Clepper has built a knowledge graph from two weeks of intercepted radio communications in Oceanus. Here are the key tasks to be analyzed:

**Task 1: Temporal Communication Patterns**

1.  Create visual analytics to identify daily patterns in communication timing and analyze how these patterns changed over the two-week observation period

2.  Select a specific entity and determine who influences them based on these patterns

**Task 2: Relationship Analysis**

1.  Develop visualizations to explore connections between vessels and people

2.  Identify closely associated groups

3.  Determine predominant topics for each group (e.g., environmentalism, Sailor Shift, fishing/leisure vessels)

**Task 3: Pseudonym Identification**

1.  Expand visualizations to reveal who is using pseudonyms

2.  Identify known pseudonyms ("Boss", "The Lookout") and discover additional ones

3.  Account for pseudonyms potentially used by multiple entities

4.  Demonstrate how these visualizations simplify entity identification

5.  Explain how understanding pseudonyms changes interpretation of activities

**Task 4: Nadia Conti Investigation**

1.  Use visual analytics to provide evidence regarding possible illegal activity by Nadia Conti

2.  Create a visual summary of Nadia's actions

3.  Assess whether Clepper's suspicions are justified

Each task builds on previous analysis to uncover the potential corruption surrounding the Nemo Reef closure and connections between officials, Sailor Shift's team, influential families, and The Green Guardians.

## 3.0 Setting up

### 3.1 Loading packages

| Package | Description |
|:-----------------------------------|:-----------------------------------|
| jsonlite | JSON parsing and generation |
| tidyverse | Core data-science suite (dplyr, ggplot2, etc.) |
| SmartEDA | Automated exploratory data analysis |
| tidygraph | Tidy tools for graph-data manipulation and analysis |
| ggraph | Grammar-based graph/network visualization |
| lubridate | Easy parsing and manipulation of dates/times |
| dplyr | Grammar of data manipulation (filter, mutate, etc.) |
| patchwork | Combine multiple ggplots into a unified layout |
| purrr | Functional programming tools for lists and vectors |
| scales | Scale functions for visualization (axis breaks, formatting) |
| RColorBrewer | Color palettes for maps and plots |
| plotly | Interactive web-based charts |
| igraph | Network analysis and graph algorithms |
| viridis | Color scales for scientific visualization (colorblind-friendly) |
| forcats | Tools for working with categorical variables (factors) |
| knitr | Dynamic report generation and literate programming |

```{r}
pacman::p_load(
  jsonlite,
  tidyverse,
  SmartEDA,
  tidygraph,
  ggraph,
  lubridate,
  dplyr,
  patchwork,
  purrr,
  scales,
  RColorBrewer,
  plotly,
  igraph,
  viridis,
  forcats,
  knitr,
  visNetwork,
  htmltools)
```

### 3.2 Loading data

```{r}
MC3_graph <- fromJSON("data/MC3_graph.json")
glimpse(MC3_graph)
```

```{r}
MC3_schema <- fromJSON("data/MC3_schema.json") # didint load 
glimpse(MC3_schema)
```

### 3.3 Examining the data structure

```{r}
nodes_tbl <- as_tibble(MC3_graph$nodes)
edges_tbl <- as_tibble(MC3_graph$edges)
```

```{r}
kable(head(nodes_tbl, 10))
kable(head(edges_tbl, 10))


```

after slpit to nodes and edges, the data shouw 3226 edges and 1159 nodes

## 4.Initial EDA

It is time for us to apply appropriate EDA methods to examine the data.

In the code chunk below, ExpCatViz() of SmartEDA package is used to reveal the frequency distribution of all categorical fields in nodes_tbl tibble dataframe.

```{r}
ExpCatViz(data = nodes_tbl,
          col = "red")
```

On the other hands, code chunk below uses ExpCATViz() of SmartEDA package to reveal the frequency distribution of all categorical fields in mc3_edges tibble dataframe.

```{r}
ExpCatViz(data=edges_tbl,
          col="lightblue")


```

## 5.0 Task 1-Clepper found that messages frequently came in at around the same time each day.

### Task 1.1: Create visual analytics to identify daily patterns in communication timing 

```{r}
ggplot(data=nodes_tbl,
      aes(y=`type`))+
  geom_bar()


```

```{r}
ggplot(data=nodes_tbl,
      aes(y=`sub_type`))+
  geom_bar()

```

**Generate Node Sub-Type Inventory**

```{r}

# Extract unique sub_types for each node type
entities      <- sort(unique(nodes_tbl$sub_type[nodes_tbl$type == "Entity"]))
events        <- sort(unique(nodes_tbl$sub_type[nodes_tbl$type == "Event"]))
relationships <- sort(unique(nodes_tbl$sub_type[nodes_tbl$type == "Relationship"]))

node_inventory <- list(
  Entities      = entities,
  Events        = events,
  Relationships = relationships
)


```

**List down type and relevant sub types**

```{r}
for (grp in names(node_inventory)) {
  cat("**", grp, "** (", length(node_inventory[[grp]]), "):\n", sep = "")
  cat(paste0("  - ", node_inventory[[grp]]), sep = "\n")
  cat("\n")
}



```

based on the above list, we noticed that based on the task requirement, now we only focus on type="event" and sub_type ="communication" , next step is to create communication nodes

```{r}
comm_nodes <-nodes_tbl %>%
  filter(type=="Event",
         sub_type == "Communication") %>%
  select(event_id = id,timestamp)

glimpse(comm_nodes)
```

**Define sender and receiver**

```{r}
ggplot(data=edges_tbl,
      aes(y=`type`))+
  geom_bar()

```

**extract data from sender and receiver**

```{r}
sent_edges <-edges_tbl %>% filter(type=="sent")%>%
  select(sender = source,event_id=target)

recv_edges <-edges_tbl %>% filter(type=="received")%>%
  select(event_id = source,receiver=target)
```

**using inner_join to combine sender and receiver**

```{r}
msgs <- sent_edges %>%
  inner_join(recv_edges, by="event_id") %>%
  inner_join(comm_nodes, by="event_id") %>%
  mutate(ts = ymd_hms(timestamp, tz="UTC")) %>%
  mutate(date = as_date(ts), hour = hour(ts))

```

**visualizing the heat map**

```{r}
heatmap_data <- msgs %>%
  count(date, hour) %>%
  complete(
    date = seq(min(date), max(date), by = "1 day"),
    hour = 0:23,
    fill = list(n = 0)
  )

ggplot(heatmap_data, aes(x=hour, y=date, fill=n)) +
  geom_tile(color="white") +
  scale_x_continuous(
    breaks = 0:23,
    expand = c(0,0)
  )+
  scale_y_date(
    breaks = seq(min(heatmap_data$date), max(heatmap_data$date), by = "1 day"),
    date_labels = "%b %d",
    expand = c(0, 0)
  ) +
 scale_fill_distiller(
    name    = "Messages",
    palette = "Spectral",
    direction = 1
  ) +
  
  labs(
    title = "Daily Communication Patterns",
    x     = "Hour of Day",
    y     = "Date"
  ) +
  theme_minimal() +
  theme(
    axis.text.x  = element_text(hjust = 0.5),
    axis.text.y  = element_text(size = 8),
    panel.grid  = element_blank()
  )


```

### Task 1.2 Compare the shape of daily rhythm between week 1 and week 2

```{r}
 # Mark Week 1 vs Week 2
msgs <- msgs %>%
  mutate(
    week = if_else(date <= min(date) + days(6), "Week 1", "Week 2")
  )

# Count and normalize within each week
week_patterns <- msgs %>%
  count(week, hour) %>%
  group_by(week) %>%
  mutate(proportion = n / sum(n))


hour_breaks <- seq(min(week_patterns$hour), max(week_patterns$hour), by = 1)
prop_breaks <- seq(
  0,
  ceiling(max(week_patterns$proportion) * 100) / 100,
  by = 0.02
)

ggplot(week_patterns, aes(x = hour, y = proportion, color = week)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_x_continuous(breaks = hour_breaks) +
  scale_y_continuous(
    breaks = prop_breaks,
    labels = percent_format(accuracy = 1)
  ) +
  labs(
    title = "Hourly Communication Patterns: Week 1 vs Week 2",
    x     = "Hour of Day",
    y     = "Percent of Total Messages"
  ) +
  theme_light(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 16, hjust = 0.5),
    legend.position = "top",
    legend.title    = element_blank(),
    panel.grid.major = element_line(color = "grey80"),
    panel.grid.minor = element_line(color = "grey90"),
    axis.text.x     = element_text(vjust = 0.5),
    axis.text.y     = element_text(size = 8)
  )



```

### Task 1.3: Focus on a specific entity and use this information to determine who has influence over them.

```{r}
entity_info <- nodes_tbl %>%
  filter(type == "Entity") %>%
  select(id, target_subtype = sub_type)

msgs2 <- msgs %>%
  left_join(entity_info, by = c("receiver" = "id"))


top_inf <- msgs2 %>%
  filter(!is.na(receiver)) %>%
  count(target_subtype, receiver, sender, sort = TRUE) %>%
  group_by(target_subtype, receiver) %>%
  slice_max(n, n = 10) %>%
  ungroup()

plot_list <- top_inf %>%
  group_split(target_subtype) %>%
  map(~{
    subtype <- unique(.x$target_subtype)
    ggplot(.x, aes(
        x     = reorder(sender, n, .desc=TRUE),
        y     = n,
        fill  = receiver,
        text  = paste0(
          "Subtype: ", subtype, "<br>",
          "Sender: ", sender, "<br>",
          "Receiver: ", receiver, "<br>",
          "Count: ", n
        )
      )) +
      geom_col(color = "white") +                    
      scale_fill_viridis_d(option = "turbo", name="Receiver") +
      coord_flip() +
      labs(
        title = paste("Subtype:", subtype),
        x     = "Influencer (Sender)",
        y     = "Message Count"
      ) +
      theme_minimal(base_size = 12) +
      theme(
        strip.text       = element_text(face = "bold"),
        axis.text.y      = element_text(size = 8),
        panel.grid.major = element_line(color = "grey80"),
        panel.grid.minor = element_line(color = "grey90"),
        legend.position  = "bottom"
      )
  })


plots <- map(plot_list, ~ggplotly(.x, tooltip = "text"))
tagList(plots)


```

::: panel-tabset
### Subtype:Group

```{r,warning=FALSE}

ee_edges <- msgs %>% select(from = sender, to = receiver)
entity_nodes <- nodes_tbl %>%
  filter(type == "Entity") %>%
  select(id, name)

entity_graph <- tbl_graph(nodes = entity_nodes, edges = ee_edges, directed = TRUE)


target <- "Neptune"


g_ig <- as.igraph(entity_graph, strict = FALSE)


root_idx <- which(V(g_ig)$name == target)


ego_vs <- unique(c(
  root_idx,
  neighbors(g_ig, root_idx, mode = "in"),
  neighbors(g_ig, root_idx, mode = "out")
))


subg <- induced_subgraph(g_ig, vids = ego_vs)


ego_graph <- as_tbl_graph(subg)


ggraph(ego_graph, layout = "kk") +
  geom_edge_link(
    aes(label = NULL),
    arrow = arrow(length = unit(4, "mm")),
    end_cap = circle(3, 'mm'),
    color = "grey70"
  ) +
  geom_node_point(
    aes(filter = (name == target)),
    size  = 6,
    color = "firebrick"
  ) +
  geom_node_point(
    aes(filter = (name != target)),
    size  = 4,
    color = "steelblue"
  ) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  labs(title = paste("Influence Network of", target)) +
  theme_graph()
```

### Subtype: Location

```{r,warning=FALSE}

ee_edges <- msgs %>% select(from = sender, to = receiver)
entity_nodes <- nodes_tbl %>%
  filter(type == "Entity") %>%
  select(id, name)

entity_graph <- tbl_graph(nodes = entity_nodes, edges = ee_edges, directed = TRUE)


target <- "Oceanus City Council"


g_ig <- as.igraph(entity_graph, strict = FALSE)


root_idx <- which(V(g_ig)$name == target)


ego_vs <- unique(c(
  root_idx,
  neighbors(g_ig, root_idx, mode = "in"),
  neighbors(g_ig, root_idx, mode = "out")
))


subg <- induced_subgraph(g_ig, vids = ego_vs)


ego_graph <- as_tbl_graph(subg)


ggraph(ego_graph, layout = "kk") +
  geom_edge_link(
    aes(label = NULL),
    arrow = arrow(length = unit(4, "mm")),
    end_cap = circle(3, 'mm'),
    color = "grey70"
  ) +
  geom_node_point(
    aes(filter = (name == target)),
    size  = 6,
    color = "firebrick"
  ) +
  geom_node_point(
    aes(filter = (name != target)),
    size  = 4,
    color = "steelblue"
  ) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  labs(title = paste("Influence Network of", target)) +
  theme_graph()
```

### Subtype:Organization

```{r,warning=FALSE}

ee_edges <- msgs %>% select(from = sender, to = receiver)
entity_nodes <- nodes_tbl %>%
  filter(type == "Entity") %>%
  select(id, name)

entity_graph <- tbl_graph(nodes = entity_nodes, edges = ee_edges, directed = TRUE)


target <- "Green Guardians"


g_ig <- as.igraph(entity_graph, strict = FALSE)


root_idx <- which(V(g_ig)$name == target)


ego_vs <- unique(c(
  root_idx,
  neighbors(g_ig, root_idx, mode = "in"),
  neighbors(g_ig, root_idx, mode = "out")
))


subg <- induced_subgraph(g_ig, vids = ego_vs)


ego_graph <- as_tbl_graph(subg)


ggraph(ego_graph, layout = "kk") +
  geom_edge_link(
    aes(label = NULL),
    arrow = arrow(length = unit(4, "mm")),
    end_cap = circle(3, 'mm'),
    color = "grey70"
  ) +
  geom_node_point(
    aes(filter = (name == target)),
    size  = 6,
    color = "firebrick"
  ) +
  geom_node_point(
    aes(filter = (name != target)),
    size  = 4,
    color = "steelblue"
  ) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  labs(title = paste("Influence Network of", target)) +
  theme_graph()
```

### Subtype:Person

```{r,warning=FALSE}

ee_edges <- msgs %>% select(from = sender, to = receiver)
entity_nodes <- nodes_tbl %>%
  filter(type == "Entity") %>%
  select(id, name)

entity_graph <- tbl_graph(nodes = entity_nodes, edges = ee_edges, directed = TRUE)


target <- "The Intern"


g_ig <- as.igraph(entity_graph, strict = FALSE)


root_idx <- which(V(g_ig)$name == target)


ego_vs <- unique(c(
  root_idx,
  neighbors(g_ig, root_idx, mode = "in"),
  neighbors(g_ig, root_idx, mode = "out")
))


subg <- induced_subgraph(g_ig, vids = ego_vs)


ego_graph <- as_tbl_graph(subg)


ggraph(ego_graph, layout = "kk") +
  geom_edge_link(
    aes(label = NULL),
    arrow = arrow(length = unit(4, "mm")),
    end_cap = circle(3, 'mm'),
    color = "grey70"
  ) +
  geom_node_point(
    aes(filter = (name == target)),
    size  = 6,
    color = "firebrick"
  ) +
  geom_node_point(
    aes(filter = (name != target)),
    size  = 4,
    color = "steelblue"
  ) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  labs(title = paste("Influence Network of", target)) +
  theme_graph()

```

### Vessel

```{r,warning=FALSE}

ee_edges <- msgs %>% select(from = sender, to = receiver)
entity_nodes <- nodes_tbl %>%
  filter(type == "Entity") %>%
  select(id, name)

entity_graph <- tbl_graph(nodes = entity_nodes, edges = ee_edges, directed = TRUE)


target <- "Reef Guardian"


g_ig <- as.igraph(entity_graph, strict = FALSE)


root_idx <- which(V(g_ig)$name == target)


ego_vs <- unique(c(
  root_idx,
  neighbors(g_ig, root_idx, mode = "in"),
  neighbors(g_ig, root_idx, mode = "out")
))


subg <- induced_subgraph(g_ig, vids = ego_vs)


ego_graph <- as_tbl_graph(subg)


ggraph(ego_graph, layout = "kk") +
  geom_edge_link(
    aes(label = NULL),
    arrow = arrow(length = unit(4, "mm")),
    end_cap = circle(3, 'mm'),
    color = "grey70"
  ) +
  geom_node_point(
    aes(filter = (name == target)),
    size  = 6,
    color = "firebrick"
  ) +
  geom_node_point(
    aes(filter = (name != target)),
    size  = 4,
    color = "steelblue"
  ) +
  geom_node_text(aes(label = name), repel = TRUE, size = 3) +
  labs(title = paste("Influence Network of", target)) +
  theme_graph()
```
:::

## Task 2: Clepper has noticed that people often communicate with (or about) the same people or vessels, and that grouping them together may help with the investigation.

### Task 2.1 :Develop visualizations to explore connections between vessels and people

based on the request, now we have to filter people and vessel

#### Step1:Clean the nodes and edges data

clean the raw nodes and edges data to ensure there are no missing values, all IDs are unique, and types are consistent.

```{r}

mc3_nodes_cleaned <- nodes_tbl %>%
  mutate(id = as.character(id)) %>%
  filter(!is.na(id)) %>%
  distinct(id, .keep_all = TRUE) %>%
  select(-thing_collected,everything())

mc3_edges_cleaned <- edges_tbl %>%
  rename(from_id = source, to_id = target) %>%
  mutate(across(c(from_id, to_id), as.character)) %>%
  filter(from_id %in% mc3_nodes_cleaned$id, to_id %in% mc3_nodes_cleaned$id) %>%
  filter(!is.na(from_id), !is.na(to_id))


```

#### Step 2: Extract person–vessel indirect links

Identify "Person" and "Vessel" nodes, and link them through "Relationship" nodes to extract all person–relationship–vessel triples.

```{r}
people_nodes <- mc3_nodes_cleaned %>% filter(sub_type == "Person")
vessel_nodes <- mc3_nodes_cleaned %>% filter(sub_type == "Vessel")
relationship_nodes <- mc3_nodes_cleaned %>% 
  filter(grepl("^Relationship_", id))

person_relationship_edges <- mc3_edges_cleaned %>%
  filter(from_id %in% people_nodes$id & to_id %in% relationship_nodes$id)

relationship_vessel_edges <- mc3_edges_cleaned %>%
  filter(from_id %in% relationship_nodes$id & to_id %in% vessel_nodes$id)

person_to_relationship <- person_relationship_edges %>%
  select(person_id = from_id, relationship_id = to_id)
relationship_to_vessel <- relationship_vessel_edges %>%
  select(relationship_id = from_id, vessel_id = to_id)

person_vessel_links <- person_to_relationship %>%
  inner_join(relationship_to_vessel, by = "relationship_id") %>%
  select(person_id, vessel_id, relationship_id)

```

#### Step 3: Add labels and types for nodes and edges 

Add labels to each person and vessel node, and add relationship type to edges for visualization.

```{r}

person_vessel_links <- person_vessel_links %>%
  left_join(people_nodes %>% select(id, person_label = label), by = c("person_id" = "id")) %>%
  left_join(vessel_nodes %>% select(id, vessel_label = label), by = c("vessel_id" = "id"))

person_vessel_links <- person_vessel_links %>%
  mutate(relationship_type = stringr::str_extract(relationship_id, "(?<=Relationship_)[^_]+"))

edges_pv <- person_vessel_links %>%
  mutate(
    from = person_label,
    to = vessel_label
  ) %>%
  select(from, to, relationship_type)


```

#### Step 4: Prepare nodes/edges table for visualization 

Combine all nodes (person + vessel) into a nodes dataframe with type and color, and add color to the edge dataframe.

```{r}
nodes <- data.frame(
  id = unique(c(edges_pv$from, edges_pv$to)),
  label = unique(c(edges_pv$from, edges_pv$to))
) %>%
  left_join(
    bind_rows(
      people_nodes %>% select(label, type = sub_type),
      vessel_nodes %>% select(label, type = sub_type)
    ),
    by = "label"
  ) %>%
  mutate(
    shape = ifelse(type == "Person", 16, 17),
    color = ifelse(type == "Person", "#FF8888", "#44C6C6")
  ) %>%
  select(id, label, type, shape, color)

relation_types <- sort(unique(edges_pv$relationship_type))
if (length(relation_types) <= 8) {
  relation_colors <- RColorBrewer::brewer.pal(length(relation_types), "Dark2")
} else {
  relation_colors <- colorRampPalette(RColorBrewer::brewer.pal(8, "Dark2"))(length(relation_types))
}
relation_color_map <- setNames(relation_colors, relation_types)

edges <- edges_pv %>%
  mutate(
    relationship = relationship_type,
    color = relation_color_map[relationship_type]
  ) %>%
  select(from, to, relationship, color)

```

#### Step 5: Plot static network with node/edge legends

Use ggraph to plot a static network diagram. Node types (person/vessel) have different shapes and colors, and the legend on the right automatically distinguishes them. Edge colors represent relationship types.

```{r}
graph_pv <- tbl_graph(
  nodes = nodes,
  edges = edges %>% rename(label = relationship),
  directed = TRUE
)

ggraph(graph_pv, layout = 'fr') +
  geom_edge_link(
    aes(color = label),
    arrow = arrow(length = unit(3, 'mm')),
    end_cap = circle(3, 'mm'),
    width = 1.4,
    alpha = 0.85,
    show.legend = TRUE
  ) +
  geom_node_point(
    aes(shape = type, color = type),
    size = 6,
    show.legend = TRUE
  ) +
  geom_node_text(
    aes(label = label),
    repel = TRUE,
    size = 3.8,
    vjust = 1.5,
    segment.color = NA 
  ) +
  scale_shape_manual(
    values = c("Person" = 16, "Vessel" = 17),
    name = "Node Type"
  ) +
  scale_color_manual(
    values = c("Person" = "#FF8888", "Vessel" = "#44C6C6"),
    name = "Node Type"
  ) +
  scale_edge_color_manual(
    values = relation_color_map,
    name = "Relationship"
  ) +
  guides(
    color = guide_legend(override.aes = list(size = 6)),
    shape = guide_legend(override.aes = list(size = 6)),
    edge_color = guide_legend(override.aes = list(linetype = 1, size = 2))
  ) +
  theme_void() +
  labs(title = "People–Vessel Interaction Network") +
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, size = 15, face = "bold")
  )

 

```


### Task 2.2 ：Identify closely associated groups and the predominant topics for each group

#### Step 1: Data preparation

referring to task 2.1 we already get - nodes: id, label, type - edges: from, to, relationship_type

```{r}
ig_undir <- as.undirected(as.igraph(graph_pv), mode = "collapse")

com <- cluster_louvain(ig_undir)
membership_vec <- membership(com) 
graph_pv <- graph_pv %>% 
  activate(nodes) %>%
  mutate(community = as.factor(membership_vec[seq_len(n())]))

nodes <- as_tibble(graph_pv, "nodes")


table(nodes$community)

```

#### Step 2: Building un-directed graph

```{r}

group_summary <- nodes %>%
  group_by(community, type) %>%
  summarise(
    members = paste(label, collapse = ", "),
    .groups = "drop"
  )
print(group_summary)


```

#### Step 3:Louvain Community detection

```{r}
library(ggraph)
library(RColorBrewer)

# 1. 社区颜色设定
n_communities <- length(unique(na.omit(nodes$community)))
# 如果社区数>8，自动扩展色盘
if(n_communities > 8) {
  palette <- colorRampPalette(brewer.pal(8, "Set2"))(n_communities)
} else {
  palette <- brewer.pal(max(3, n_communities), "Set2")
}

# 2. 边(relationship)颜色设定，自动取 Dark2 色盘
relation_types <- unique(edges$label)
n_rel <- length(relation_types)
relation_colors <- brewer.pal(max(3, n_rel), "Dark2")
relation_color_map <- setNames(relation_colors, relation_types)

ggraph(graph_pv, layout = 'fr') +
  geom_edge_link(
    aes(color = label),
    arrow = arrow(length = unit(3, 'mm')), 
    end_cap = circle(3, 'mm'),
    width = 1.3,
    alpha = 0.9,
    show.legend = TRUE
  ) +
  geom_node_point(
    aes(shape = type, fill = community),
    size = 9,
    color = "black",
    stroke = 1.1,
    alpha = 1,
    show.legend = TRUE
  ) +
  geom_node_text(
    aes(label = label),
    repel = TRUE,
    size = 3,
    segment.color = NA
  ) +
  scale_shape_manual(
    values = c("Person" = 21, "Vessel" = 24),
    name = "Node Type"
  ) +
  scale_fill_manual(
    values = palette,
    name = "Community"
  ) +
  scale_edge_color_manual(
    values = relation_color_map,
    name = "Relationship"
  ) +
  guides(
    fill = guide_legend(
      override.aes = list(shape = 21, size = 7, color = "black", stroke = 1.2)
    ),
    shape = guide_legend(
      override.aes = list(size = 7, fill = "#CCCCCC", color = "black", stroke = 1.2)
    ),
    edge_color = guide_legend(
      override.aes = list(linetype = 1, size = 3, alpha = 1)
    )
  ) +
  theme_void() +
  labs(title = "People–Vessel Network: Community Structure") +
  theme(
    legend.position = "right",
    legend.box = "vertical",
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
  )
  



```


```{r}
group_summary$topic <- case_when(
  grepl("EcoVigil|Green Guardians", group_summary$members) ~ "Environmentalism",
  grepl("Sailor Shift|Mariner's Dream", group_summary$members) ~ "Entertainment",
  grepl("Mako|Remora|Marlin|Neptune", group_summary$members) ~ "Fishing/Leisure",
  TRUE ~ "Other"
)
print(group_summary)


```

